### 2024-07-11 src/main.tex
changed to ACM format

\documentclass[manuscript, screen, acmsmall]{acmart}
\usepackage[noend]{algorithm2e}
\usepackage{graphicx,balance,multirow,xspace,subcaption,longtable}
\usepackage{tabularx,multirow,booktabs,blindtext}
\usepackage{makecell,amsmath,cleveref,rotating}
\usepackage[font={small}, textfont=md]{caption}
\usepackage[utf8]{inputenc}
\usepackage[draft,inline,nomargin,index]{fixme}
\usepackage{marginnote}
\usepackage[bb=boondox,bbscaled=.95,cal=boondoxo]{mathalfa}
\usepackage{adjustbox}
\usepackage{colortbl}

\settopmatter{printacmref=false}
\setcopyright{none}

\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

\PassOptionsToPackage{hyphens}{url}
\usepackage{hyperref} \urlstyle{tt}
\newcolumntype{P}[1]{>{\arraybackslash}p{#1}}
%\newcolumntype{X}[1]{>{\centering\arraybackslash}p{#1}}

\crefformat{section}{\S#2#1#3}
\crefformat{subsection}{\S#2#1#3}
\crefformat{subsubsection}{\S#2#1#3}

\hypersetup{%
  pdftitle={Report for Qualifying Exam},
  pdfauthor={},
  pdfkeywords={},
  bookmarksnumbered,
  pdfstartview={FitH},
  colorlinks,
  citecolor=black,
  filecolor=black,
  linkcolor=black,
  urlcolor=linkColor,
  breaklinks=true,
  hypertexnames=false
}

\clubpenalty=10000
\widowpenalty=10000

\newcommand{\red}[1]{\textcolor{red}{\bf #1}}
\newcommand{\blue}[1]{\textcolor{blue}{\bf #1}}

\newenvironment{packed_itemize}{
\begin{itemize}
  \setlength{\itemsep}{2pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
  \setlength{\topsep}{2pt}
}{\end{itemize}}

\newenvironment{packed_enumerate}{
\begin{enumerate}
  \setlength{\itemsep}{2pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
  \setlength{\topsep}{2pt}
}{\end{enumerate}}

% \renewcommand{\footnotesize}{\scriptsize}
\newcommand{\etc}{etc.}
\newcommand{\eg}{e.g.,\ }
\newcommand{\etal}{et al.\xspace}
\newcommand{\ie}{i.e.,\ }
\newcommand{\re}{r.e.\ }
\newcommand{\aka}{a.k.a.\ }
\newcommand{\cf}{\emph{cf.\ }}
\newcommand{\note}[1]{\textcolor{blue}{#1}}
\newcommand{\superem}[1]{\underline{\emph{#1}}}
\newcommand{\para}[1]{\vspace{.05in}\noindent\textbf{#1}}
\newcommand{\paraem}[1]{\vspace{.05in}\noindent\emph{#1}}
\newcommand{\yt}{YouTube\xspace}
\newcommand{\X}{$\mathbb{X}$\xspace}
\newcommand\mycommfont[1]{\footnotesize\ttfamily\textcolor{blue}{#1}}

\begin{document}

\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\pagestyle{plain} % removes running headers

\renewcommand{\sectionautorefname}{\S}
\renewcommand{\subsectionautorefname}{\S}
\renewcommand{\subsubsectionautorefname}{\S}

\input{sections/titlepage}
%%%%%%%%%%%%%%%%%%%%%%%%%%% Table of Contents
\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document
%\maketitle
%\sloppy
\input{sections/abstract}
\input{sections/introduction}
\newpage
\input{sections/relatedwork}
\newpage
\input{sections/papers}
\balance

\newpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{main}

\end{document}