### 2024-07-11 src/sections/papers.tex

%! Author = mkeim
%! Date = 7/11/24

\section{Research Papers} \label{sec:paper}
\subsection{Statistically Significant Detection of Linguistic Change}\label{subsec:statistically-significant-detection-of-linguistic-change}

\subsubsection{About}
This research aims to quantify linguistic shifts in word meaning and usage across time, focusing on semantic or contextual changes.
The study utilizes diverse datasets, including Twitter posts, Amazon product reviews, and Google Book Ngrams, to construct time series for individual words.
Three distinct methods are employed for statistical modeling:
\begin{itemize}
    \item frequency analysis to capture sudden changes in word usage,
    \item syntactical analysis examining each word's part-of-speech tag distribution, and
    \item distributional analysis investigating contextual cues through word co-occurrence statistics.
\end{itemize}
To determine the statistical significance of word changes over time, the authors are developing a change point detection algorithm.
Additionally, the research incorporates time series data representing the percentage of search queries according to Google Trends, providing further insight into evolving language patterns and word usage shifts.

\subsubsection{Research Questions}
\begin{enumerate}
\item What methods can quantify the statistical relevance of observed changes in a word's usage across different time periods?
\item Given that a word's usage has changed, how can the precise moment or period of this shift be determined?
\end{enumerate}

\subsubsection{Methodology}

\subsubsection{Results}

\subsection{Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change}\label{subsec:diachronic-word-embeddings-reveal-statistical-laws-of-semantic-change}
\subsubsection{About}
\subsubsection{Research Questions}
\subsubsection{Methodology}
\subsubsection{Results}

\subsection{Specious Sites: Tracking the Spread and Sway of Spurious News Stories at Scale}\label{subsec:specious-sites:-tracking-the-spread-and-sway-of-spurious-news-stories-at-scale}
\subsubsection{About}
\subsubsection{Research Questions}
\subsubsection{Methodology}
\subsubsection{Results}