### 2024-08-19 src/sections/introduction.tex
changed introduction and minor changes

%! Author = mkeim
%! Date = 7/11/24

\section{Introduction} \label{sec:introduction}
Language is a dynamic, living entity that continually adapts to reflect the changing world around us.
As language evolves, words shift in meaning, acquire new connotations, or are repurposed for entirely new contexts.
These changes occur as society, technology, and culture change.
When words change meaning, there is often a delay before search engines and online platforms fully reflect these new meanings.
This creates a \emph{data void} — an absence of relevant, up-to-date information based on the current understanding of the word.
This void represents a gap between evolving language and existing online content.

\para{Research goal.}
Understanding how words evolve is crucial because it allows us to anticipate where data voids might occur.
By tracking linguistic changes, our goal is to predict when a word might create a gap in digital information which could lead to a potential data void.
To achieve this, we have selected following research papers as a foundation for our research.

\begin{itemize}
    \item Statistically Significant Detection of Linguistic Change.
    ACM WWW 2015 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic} (\Cref{sec:paper_kulkarni})
    \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change.
    ACL 2016 ~\cite{hamilton-etal-2016-diachronic} (\Cref{sec:paper_hamilton})
\end{itemize}

The concept of semantic change, as presented in these papers, refers to the change in the contextual usage of words over time.
This can be particularly useful for analyzing how words evolve and ensuring that, as meanings shift, data voids are promptly recognized and addressed.

\para{Finding data voids is significantly important.}
Data voids have real-world implications, particularly in socio-political and sociotechnical domains.
When important terms, especially those tied to politics, social issues, or technology, develop new meanings,
data voids can be exploited to shape public perception.

Data voids, left unaddressed, can distort debates, influence elections, and shape collective beliefs, often in misleading or harmful directions.
Once a data void is identified, it’s critical to fill that gap with accurate, authoritative information.
Otherwise, bad actors can capitalize on the absence of reliable data to promote their own agenda.
By understanding linguistic changes, researchers and technology platforms can work together to prevent these voids from becoming filled with harmful or misleading content.

\para{Example.}
The term \emph{crisis actor} has undergone a significant shift in meaning,
exemplifying the phenomenon of semantic change we've been discussing.

Historically, `crisis actor' referred to actors employed in emergency preparedness training exercises.
These actors simulate victims or other roles in staged disaster drills to help first responders practice their response to crises like natural disasters or terrorist attacks.
It was a technical, niche term with a specific and legitimate use.

However, in recent years, it is used to describe the idea that victims of real-world tragedies are actually actors.
For example, in mass shootings, the victims have been referred to in propaganda and misinformation campaigns as `crisis actors,'
falsely implying that the events were staged or fabricated.

This example clearly demonstrates how a data void occurs when a word’s meaning is co-opted and evolves rapidly,
leaving a gap in reliable, accurate content.
This allows misinformation to flourish, having a profound impact on both political discourse and social trust.\\

This report introduces methodologies to identify shifts in words and this will be useful in my research to proactively identify data voids based on two key papers.
To contextualize our work, we present a thorough literature review examining the evolution of word meaning (\Cref{sec:relatedwork}).
Subsequently, Section \Cref{sec:paper_kulkarni} delves into the first paper, focusing on significant linguistic change detection,
while Section \Cref{sec:paper_hamilton} explores the second paper on diachronic word embeddings.